<!--
---
title: "How to Place a Bet with abettor"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to Place a Bet with abettor}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---
-->

# How to Place a Bet with abettor

## Purpose

The purpose of this document is to demonstrate how to place a bet on the Betfair betting exchange, using their API-NG product, the `abettor` R package and a simple R script.

Many of the steps outlined in this tutorial will be purposefully simplistic, in order to demonstrate a wide range of `abettor` functions. It is assumed that readers have no prior experience using any Betfiar API products.

## Scope

This tutorial will outline a workflow for placing a simple Win Only bet on a horse racing market.

It is assumed that the bettor knows only the following three pieces of information:

* Horse Name
* Race Course Name
* Race Start Time

Two further pieces of information must be obtained from Betfair:

* Market ID
* Selection ID

An indication of the best price available and the market depth will also be demonstrated, although this is not strictly necessary for placing a bet.

Using just these few simple pieces of information it is possible to use the `abettor` package to place a bet.

## Tutorial

### Load Required Packages

Only three packages or libraries are required. All packages obviously must be installed first.

```{r, eval = FALSE}
require("RCurl")
require("jsonlite")
require("abettor")
```

### Login to Betfair

```{r, eval = FALSE}
loginBF(username = "YourBFUsername", password = "YourBFPassword", applicationKey = "YourAppKey")
```

Application keys are straight forward to obtain. In short:

* Login to the Betfair website
* Open a new browser tab and go the Betfair [Accounts API Visualiser](https://developer.betfair.com/visualisers/api-ng-account-operations/).
* The session token field at the top of the page should be pre-filled. Select createDevelopAppKeys in the left column.
* Enter an Application Name. Don't enter something silly. Each user is only allowed one set of Application Keys, so once an application name is chosen, you're stuck with it. (If you receive a vague error, the chances are you already have an application key. Therefore, select getDeveloperAppKeys in the left hand column and proceed.)
* Click Execute at the bottom of the page.
* Your application keys will be shown in the right hand column. Use the DELAY key for testing and non-price related functions. You will need to use the non-DELAY application key to complete this tutorial and place a bet.

Betfair provides further [details here](https://api.developer.betfair.com/services/webapps/docs/display/1smk3cen4v3lu3yomq5qye0ni/Application+Keys).

After logging in, if all is well, you should not have received an error. There is no success message displayed by default by `abettor`. If you wish to check, you should look at the `headersPostLogin` variable. This will contain not only your application key, but also a relevant session token. Try `headersPostLogin$'X-Authentication'`

### Find Relevant Event Type

Which sport? We need to know the relevant event type identifaction number for our chosen sport, in this case Horse Racing.

```{r, eval = FALSE}
listEventTypes()
```

This will return a list of all sports with markets on Betfair. Horse racing has an event type ID of 7.

### Find Relevant Country Code

This one is pretty straightforward. If you know your event venue, then you'll know which country it is in, but you will need to know the specific two letter country code.

```{r, eval = FALSE}
listCountries(eventTypeIds = "7")
```

As you can see, we've passed an argument to the function, effectively limiting results to just countries with current horse racing markets.

More information regarding any function within the abettor package, and indeed any R package, may be found like this:

```{r, eval = FALSE}
?listCountries
```

### Find Relevant Market Type

What type of bet do you wish to place? Win, Place or antepost perhaps?

```{r, eval = FALSE}
listMarketTypes(eventTypeIds = "7")
```

The returned results show just the market types available for Horse Racing.

### Find Market and Selection IDs

Everything we've done so far has been very straightforward and easily achieved with a single `abettor` function and a couple of arguments. Now we start the real work in finding the relevant details for our bet. 

Both the Market ID and Selection ID can be found through a single `abettor` function, but we'll need a little more R code to extract them.

```{r, eval = FALSE}
marketCat <- listMarketCatalogue(eventTypeIds = "7", marketCountries = "GB", marketTypeCodes = "WIN")
```

We've used the `listMarketCatalogue` function, limiting results to Horse Racing in Great Britain and Win only markets. By default `listMarketCatalogue` returns results for markets starting immediately and 24 hours into the future. These default times may be changed. `?listMarketCatalogue` for more details. We've also stored the returned data in a variable called `marketCat`.

Upon examining `marketCat` variable you can immediately see it containts a lot of data. This data does include the Market ID and Selection ID we require. These can be viewed in isolation with the following:

```{r, eval = FALSE}
marketCat$marketId
marketCat$runners[[1]]$selectionId
```

Have a really good look at `marketCat`, it contains a lot of data and is quite complex with nested lists and data frames.

```{r, eval = FALSE}
head(marketCat)
str(marketCat)
```

As we already know a couple of things about the bet we wish to place, including the course name, start time and the horse name, we can now subset `marketCat` to return a much narrower and relevant set of data. There's about eight million ways of subsetting data in R, this is just one of them.

Subset using the known course name and race start time. In this case we have a sure thing in the last race on the Wolverhampton card at 17:40. Previously you will have taken note of the Betfair specific time format, such as that found in `marketCat$marketStartTime`

```{r, eval = FALSE}
ourSpecificRace <- marketCat[grep("^Wolverhampton", marketCat$event$venue), ]
ourSpecificRace <- ourSpecificRace[grep("^2014-12-15T17:40*", ourSpecificRace$marketStartTime),]
```

Here's our Market ID.

```{r, eval = FALSE}
ourMarketId <- ourSpecificRace$marketId
```

Retain just the fields we want. `ourSpecificRace$runners` is a list with a nested data frame. The data we want is in the first element of that list.

```{r, eval = FALSE}
runners <- ourSpecificRace$runners[[1]][ , c("selectionId","runnerName")]
```

Find the `selectionId` of just the horse we wish to bet on, a champion galloper called Memoria.

```{r, eval = FALSE}
ourRunner <- runners[grep("^Memoria", runners$runnerName), ]
ourSelectionId <- ourRunner$selectionId
```

Check we have sensible looking data by examining the saved variable values.

```{r, eval = FALSE}
ourMarketId
ourSelectionId
```

### Find Available Prices and Market Depth

We're nearly there and have all the essential elements we need to place a bet. The last thing it would be useful to know is what prices are available for our chosen selection. `listMarketBook` supplies that information. Again we're dealing with nested lists and data frames.

```{r, eval = FALSE}
ourMarketIdPrices <- listMarketBook(marketIds = ourMarketId, priceData = "EX_ALL_OFFERS")
allRunnersPrices <- ourMarketIdPrices$runners
ourSelectionIdPrices <- allRunnersPrices[[1]][which(allRunnersPrices[[1]]$selectionId == ourSelectionId),]
```

We're interested in backing to Win, so creating a final data frame with prices available and market depth can be obtained.

```{r, eval = FALSE}
ourSelectionIdPricesDF <- data.frame(ourSelectionIdPrices$ex[[1]])
```

Create a variable to store the best price. This is not really needed, but may be handy later.

```{r, eval = FALSE}
ourSelectionBestPrice <- ourSelectionIdPricesDF[1,c("price")]
```

At the time of writing, the best price available on Memoria was 6.4 with £8.33 available to be matched.

### Place A Bet

Finally, we'll place a £2 bet using the `placeOrders` function.

```{r, eval = FALSE}
PlaceBetReturn <- placeOrders(marketId = ourMarketId, selectionId = ourSelectionId, betSide = "BACK", betType = "LIMIT", betSize = "2", reqPrice = ourSelectionBestPrice, persistenceType = "LAPSE")
```

If everything has been successful, the `PlaceBetReturn` variable will contain a unique customer reference, which defaults to the exact date and time the bet was placed. See `?placeOrders` to alter this.

We've now reached the end of this tutorial and hopefully you've successfully placed a bet on Betfair using R.

## Final Words

If you ever need to know more about `abettor` functions from within R, simply type `?functionName`.

It's probably a reasonable idea to consult the [Betfair API-NG developer documentation](https://api.developer.betfair.com/services/webapps/docs/display/1smk3cen4v3lu3yomq5qye0ni/API-NG+Overview), but it's by no means essential in order to successfully use `abettor`

---


