---
title: "How to extract runner data to a dataframe with abettor"
author: "Soccerama"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extracting runner data to a dataframe}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# How to extract runner data to a dataframe with abettor

## Purpose

The purpose of this document is to provide an example of how to extract runner data from the Betfair betting exchange to a dataframe using their API-NG product, the `abettor` R package and a simple R script. This runner data can then be used to place bets using the API.

This assumes that you have read the vignette on placing a simple bet. To avoid duplication, this vignette does not cover:
* logging in
* getting an Application Key

## Scope

This tutorial will outline a workflow for extracting runner data for a horse racing market.

It is assumed that the bettor wants the runner data for the race given:

* Race Course Name
* Race Start Time

We use this information to obtained the market ID from the Betfair API. Runner data is then obtained for each horse in the race.

## Tutorial

### Prior reading

This tutorial does not assumes you have read the [initial tutorial](https://github.com/phillc73/abettor/blob/master/vignettes/abettor-placeBet.Rmd) which contains information on obtaining and Application Key and logging in.

### Load Required Packages

Only two packages or libraries are required. All packages obviously must be installed first.

```{r, eval = FALSE}
# Requires a minimum of version 1.7.2
# Older versions will most likely work but are unsupported
library("jsonlite")
# Requires a minimum of version 1.4.2
# Older versions will most likely work but are unsupported
library("httr")
require("abettor")
```

### Login to Betfair

```{r, eval = FALSE}
loggedin <- loginBF(username = "YourBFUsername", password = "YourBFPassword", applicationKey = "YourAppKey")
```

You will need an Application Key and the guide to obtaining one is in the [initial tutorial](https://github.com/phillc73/abettor/blob/master/vignettes/abettor-placeBet.Rmd)

Betfair provides further [details here](https://docs.developer.betfair.com/display/1smk3cen4v3lu3yomq5qye0ni/Application+Keys).

You can check that the log in was successful by looking at the value of `loggedin` which should be `"SUCCESS:"`

### Choose event

Following on from the initial tutorial, we look at a win market for a horse race in Great Britain. We can easily get a list of events by using the function.

```{r, eval = FALSE}
marketCat <- listMarketCatalogue(eventTypeIds = "7", marketCountries = "GB", marketTypeCodes = "WIN")
```

This returns a list of all horse racing win markets in Great Britain starting in the next 24 hours which is stored in a dataframe called marketCat. There is a lot of data in this dataframe and we don't need all of that for now. We're looking for a specific race and we can review the relevant information for each race using:

```{r, eval = FALSE}
cbind(marketCat$event$venue,marketCat$marketName,marketCat$marketStartTime,marketCat$marketId)
```

The starting time shown here is in UTC so you may want to check you have the correct event. The initial tutorial goes through how you can subset a race using R. You can either use that code or identify the marketId visually and store it in `myMarketId`.

To simplify things, we want just our market which we can get using:
```{r, eval=FALSE}
mc <- marketCat[marketCat$marketId == myMarketId,]
```

### Get exchange information for your chosen market

To get information on the current exchange position there is another function called `listMarketBook`. 
```{r, eval=FALSE}
winOdds <- listMarketBook(marketId = myMarketId, priceData = "EX_BEST_OFFERS")
```

The information about the current market position is in the `runners` dataframe but there are some other useful items here worth noting. The `status` field gives the current event status and you may want to use this as a check in your code. There is also an `inplay` field which may be helpful. Fields `numberOfRunners` and `numberOfActiveRunners` show you if there have been anyscratchings and how many horses are currently expected to run.

There is more information that can be retrieved using `listMarketBook` using the `priceProjection`, `orderProjetion` and `matchProjection` flags and you can adapt this code to get exactly the data you want.

The market information for each horse is in the `runners` dataframe which you can access with

```{r, eval=FALSE}
winOdds$runners
```

If there have been any scratchings in the race since the field was posted by Betfair, they will still appear in the dataframe with the `status` field equal to `REMOVED`. If a horse was scratched before the field was posted by Betfair it won't appear.

The code below extracts the back and lay offers and puts the best price and size for each of these into a dataframe. Any missing entries will be left as `NA`.

```{r, eval=FALSE}
numRunners <- winOdds$numberOfRunners
if(winOdds$status == "OPEN"){
  runnerData <- data.frame(runnerName = mc$runners[[1]][,"runnerName"])
  allRunnerData <- winOdds$runners[[1]]
  runnerData[,"selectionId"] <- allRunnerData$selectionId
  for(j in 1:numRunners){
    if(allRunnerData$status[[j]]=="ACTIVE"){
      #Check that there are orders in the stack for this runner or it causes an error
      if(nrow(allRunnerData$ex[[1]][[j]])>0){
        runnerData[j,"bestLayPrice"] <- allRunnerData$ex[[1]][[j]][[1,"price"]]
        runnerData[j,"layAmount"] <- allRunnerData$ex[[1]][[j]][[1,"size"]]
      } 
      if(nrow(winOdds$runners[[1]]$ex[[2]][[j]])>0){
        runnerData[j,"bestBackPrice"] <- allRunnerData$ex[[2]][[j]][[1,"price"]]
        runnerData[j,"backAmount"] <- allRunnerData$ex[[2]][[j]][[1,"size"]]
      } 
    } 
  }
}
```

### What next?

This tutorial shows how you can extract the current best market price for each runner to feed into your bot. There are more variations of what you can get from `listMarketBook` including information on BSP. You may want to read the documentation for the [listMarketBook function](https://docs.developer.betfair.com/display/1smk3cen4v3lu3yomq5qye0ni/listMarketBook) and explore other data that is available.

Your bot can now use this information to place a bet. If you need help with this step then you should read the initial tutorial on [placing your first bet]( https://github.com/phillc73/abettor/blob/master/vignettes/abettor-placeBet.Rmd)

## Final Words

If you ever need to know more about `abettor` functions from within R, simply type `?functionName`.

The [Betfair API-NG developer documentation](https://api.developer.betfair.com/services/webapps/docs/display/1smk3cen4v3lu3yomq5qye0ni/API+Overview) also contains a wealth of information and answers to your questions.

---
